---
title: "Machine Learning Project"
---

```{r warning = FALSE, message = FALSE}
# Suppress dplyr summarise grouping warning messages
options(dplyr.summarise.inform = FALSE)

library(tidyverse)
library(tidymodels)

credit_card_df <- readRDS(url('https://gmubusinessanalytics.netlify.app/data/credit_card_df.rds'))

```

# Data Analysis

In this section, you must think of at least 5 relevant questions that explore the relationship between `customer_status` and the other variables in the `credit_card_df` data set. The goal of your analysis should be discovering which variables drive the differences between customers who do and do not close their account.

You must answer each question and provide supporting data summaries with either a summary data frame (using `dplyr`/`tidyr`) or a plot (using `ggplot`) or both.

In total, you must have a minimum of 3 plots (created with `ggplot`) and 3 summary data frames (created with `dplyr`) for the exploratory data analysis section. Among the plots you produce, you must have at least 3 different types (ex. box plot, bar chart, histogram, scatter plot, etc...)

See the [Data Analysis Project](https://gmubusinessanalytics.netlify.app/data-analysis-project.html){target="_blank"} for an example of a question answered with a summary table and plot.

**Note**: To add an R code chunk to any section of your project, you can use the keyboard shortcut `Ctrl` + `Alt` + `i` or the `insert` button at the top of your R project template notebook file.



# Question 1


**Question**: Customers belonging to which employment status have higher closed and active account status?

**Answer**:
From the bar graph, we can see that most credit card accounts have been closed by part-time employees. Uncertainty of job/income might be the reason part-time employees might have closed their credit card accounts. Full-time employees have the maximum count who has active credit card account, stable job and income might be the reason for active customer status.

```{r}
credit_card_df %>% group_by(customer_status,employment_status) %>% 
                  summarise( count = n())

ggplot(credit_card_df, aes(employment_status , ..count..)) + geom_bar(aes(fill = customer_status), position = "dodge") 

```



# Question 2


**Question**: Identify if there is any relation between marital_status and customer status?


**Answer**:
From this analysis, customers with divorced marital status have the least count of 195 customers who have active credit card accounts as compared to single and married customers. It is also seen that customers with divorced marital status have the least customer count of 159 who has closed the credit card account. Hence, there is no relation between marital status and customer status.

```{r}

credit_card_df %>% group_by(customer_status,marital_status) %>% 
                  summarise(n_count = n())

```


# Question 3


**Question**: What is the relation between customer_status and utilization_ratio?


**Answer**:
From the histogram, we observe active customers have high utilization ratio than the customers who have closed their accounts. This might indicate that customers who are not using a credit card are willing to close their accounts.


```{r}

credit_card_df %>% group_by(customer_status) %>% 
                  summarise(n_utilization_ratio = n(),
                            min_utilization_ratio = min(utilization_ratio),
                            max_utilization_ratio = max(utilization_ratio),
                            avg_utilization_ratio = mean(utilization_ratio))
                            
ggplot(data = credit_card_df, aes(x = utilization_ratio, fill = customer_status)) +geom_histogram(aes(y = ..density..), color = "white", bins = 15) +
   facet_wrap(~ customer_status, nrow = 2) +
   labs(title = "Customer Status Of Accounts",x = "utilization_ratio", y = "count") 

```



# Question 4


**Question**: Does education affect the customer status ?


**Answer**:
The following analysis shows that people having education as associates and masters are the highest who have closed credit card accounts among all others. Similarly, associates and masters are also the highest among all who have the maximum active account status. Hence we can infer that education does not have a significant effect on customer status.

```{r}
credit_card_df %>% group_by(customer_status,education) %>% 
                  summarise(n_count = n())

```



# Question 5


**Question**: Identify the relation between the customer_status and income?


**Answer**: From the below analysis, we can see that median of active customers is slightly more than the median of the customers who have closed their accounts. Hence, we cannot clearly say that customers which have high salaries have active credit accounts because their median differences are negligible.

```{r}
ggplot(credit_card_df, aes(x = customer_status, y = income)) +
        geom_boxplot() 

```

# Machine Learning


In this section of the project, you will fit **three classification algorithms** to predict the outcome variable,`customer_status`.

You must follow the machine learning steps below. 

The data splitting and feature engineering steps should only be done once so that your models are using the same data and feature engineering steps for training.

- Split the `credit_card_df` data into a training and test set (remember to set your seed)
- Specify a feature engineering pipeline with the `recipes` package
    - You can include steps such as skewness transformation, correlation filters, dummy variable encoding or any other steps you find appropriate
- Specify a `parsnip` model object
    - You may choose from the following classification algorithms:
      - Logistic Regression
      - LDA
      - QDA
      - KNN
      - Decision Tree
      - Random Forest
- Package your recipe and model into a workflow
- Fit your workflow to the training data
    - If your model has hyperparameters:
      - Split the training data into 5 folds for 5-fold cross validation using `vfold_cv` (remember to set your seed)
      - Perform hyperparamter tuning with a random grid search using the `grid_random()` function
      - Refer to the following tutorial for an example - [Random Grid Search](https://gmubusinessanalytics.netlify.app/lesson-08-r-tutorial.html#Hyperparameter_Tuning14){target="_blank"}
      - Hyperparameter tuning can take a significant amount of computing time. Be careful not to set the `size` argument of `grid_random()` too large. I recommend `size` = 10 or smaller.
      - Select the best model with `select_best()` and finalize your workflow
- Evaluate model performance on the test set by plotting an ROC curve using `autoplot()` and calculating the area under the ROC curve on your test data

```{r}

#Data Splitting 

set.seed(250)

credit_card_split <- initial_split( credit_card_df , prop = 0.75,
                              strata = customer_status)

credit_card_training <- credit_card_split %>% training()

credit_card_test <- credit_card_split%>% testing()


#Feature Engineering 

credit_card_recipe <- recipe(customer_status ~., data = credit_card_df) %>% 
                 step_YeoJohnson(all_numeric(), -all_outcomes()) %>% 
                 step_normalize(all_numeric(), -all_outcomes()) %>% 
                 step_dummy(all_nominal(), -all_outcomes())


credit_card_recipe %>% 
  prep() %>% 
  bake(new_data = credit_card_training)


set.seed(250)
credit_card_folds <- vfold_cv(credit_card_df, v = 5)
levels(credit_card_df$customer_status)


```



# Model 1

```{r}

#Random Forest Model

library(ranger)
library(ggplot2)
random_forest_model <- rand_forest(mtry = tune(),
                        trees = tune(),
                        min_n = tune()) %>% 
            set_engine('ranger', importance = "impurity") %>% 
            set_mode('classification')

random_forest_workflow <- workflow() %>% 
               add_model(random_forest_model) %>% 
               add_recipe(credit_card_recipe)


set.seed(250)

random_forest_grid <- grid_random(mtry() %>% range_set(c(3, 5)),
                       trees(),
                       min_n(),
                       size = 5)

random_forest_grid

set.seed(250)

random_forest_tuning <- random_forest_workflow %>% 
             tune_grid(resamples = credit_card_folds,
                       grid = random_forest_grid)

random_forest_tuning %>% show_best('roc_auc')

best_random_forest <- random_forest_tuning %>% 
                      select_best(metric = 'roc_auc')


best_random_forest

final_random_forest_workflow <- random_forest_workflow %>% 
                     finalize_workflow(best_random_forest)

random_forest_wf_fit <- final_random_forest_workflow %>% 
             fit(data = credit_card_training)

random_forest_fit <- random_forest_wf_fit %>% 
          extract_fit_parsnip()
random_forest_last_fit <- final_random_forest_workflow %>% 
               last_fit(credit_card_split)



#ROC Curve
random_forest_last_fit %>% collect_predictions() %>% 
                roc_curve(truth  = customer_status, estimate = .pred_closed_account) %>% 
                autoplot()

#Confusion Matrix 
random_forest_predictions <- random_forest_last_fit %>% collect_predictions()

#Confusion Matrix 
conf_mat(random_forest_predictions, truth = customer_status, estimate = .pred_class)


# Accuracy and Fmeasure
my_metrics<- metric_set(accuracy,f_meas, recall)
my_metrics(random_forest_predictions, truth=customer_status, estimate=.pred_class)
 



```





# Model 2

```{r}

# KNN Model

library(kknn)
set.seed(250)
knn_model <- nearest_neighbor(neighbors =tune()) %>% 
             set_engine('kknn') %>% 
             set_mode('classification')


knn_wf <- workflow() %>% 
          add_model(knn_model) %>% 
          add_recipe(credit_card_recipe)


k_grid <- tibble(neighbors = c(10, 15, 25, 45, 60, 80, 100, 120, 140, 180))



set.seed(250)

knn_tuning <- knn_wf %>% 
              tune_grid(resamples = credit_card_folds, grid = k_grid)

best_k <- knn_tuning %>% 
          select_best(metric = 'roc_auc')

final_knn_wf <- knn_wf %>% 
                finalize_workflow(best_k)

knn_fit <- final_knn_wf %>% 
           last_fit(split = credit_card_split)

knn_results <-   knn_fit %>% 
                 collect_predictions()

#ROC Curve
roc_curve( knn_results , truth = customer_status , estimate = .pred_closed_account ) %>% 
  autoplot()


roc_auc(knn_results, truth = customer_status, .pred_closed_account)


#Confusion Matrix 
conf_mat(knn_results, truth = customer_status, estimate = .pred_class)


# Accuracy, Fmeasure and Recall
my_metrics<- metric_set(accuracy,f_meas,recall)
my_metrics(knn_results,truth=customer_status,estimate=.pred_class)




```





# Model 3

```{r}

#Logistic Regression 

set.seed(250)
log_reg_model <- logistic_reg() %>% 
                  set_engine('glm') %>% 
                  set_mode('classification')

log_reg_wf <- workflow() %>% 
               add_model(log_reg_model) %>% 
               add_recipe(credit_card_recipe)

log_reg_fit <-  log_reg_wf %>% 
                 last_fit(split=credit_card_split)

log_reg_results <-  log_reg_fit %>% 
                     collect_predictions()


#ROC Curve
roc_curve( log_reg_results , truth = customer_status,estimate = .pred_closed_account ) %>% 
  autoplot()


roc_auc(log_reg_results, truth = customer_status, .pred_closed_account)

#Confusion Matrix 
conf_mat(log_reg_results, truth = customer_status, estimate = .pred_class)


# Accuracy Fmeasure and Recall
my_metrics<- metric_set(accuracy,f_meas,recall)
my_metrics(log_reg_results,truth=customer_status,estimate=.pred_class)



```




# Summary of Results

Write a summary of your overall findings and recommendations to the executives at the bank. Think of this section as your closing remarks of a presentation, where you summarize your key findings, model performance, and make recommendations to improve customer retention and service at the bank.

Your executive summary must be written in a [professional tone](https://www.universalclass.com/articles/writing/business-writing/appropriate-tone-in-business-communications.htm){target="_blank"}, with minimal grammatical errors, and should include the following sections:

1. An introduction where you explain the business problem and goals of your data analysis

    - What problem(s) is this company trying to solve? Why are they important to their future success?
  
    - What was the goal of your analysis? What questions were you trying to answer and why do they matter?

<br>

2. Highlights and key findings from your Exploratory Data Analysis section 
    - What were the interesting findings from your analysis and **why are they important for the business**?

    - This section is meant to **establish the need for your recommendations** in the following section

<br>

3. Your “best” classification model and an analysis of its performance 
    - In this section you should talk about the expected error of your model on future data
      - To estimate future performance, you can use your model performance results on the **test data**
    - You should discuss at least one performance metric, such as an F1, sensitivity, specificity, or ROC AUC for your model. However, you must explain the results in an **intuitive, non-technical manner**. Your audience in this case are executives at a bank with limited knowledge of machine learning.

<br>

4. Your recommendations to the bank on how to reduce the number of customers closing their credit card accounts 
  
    - Each recommendation must be supported by your data analysis results 

    - You must clearly explain why you are making each recommendation and which results from your data analysis support this recommendation

    - You must also describe the potential business impact of your recommendation:
      
      - Why is this a good recommendation? 
      
      - What benefits will the business achieve?


**Summary**

Add your summary here. Please do not place your text within R code chunks.
>Introduction:

The main objective of this project is to determine the factors that have led the customers to cancel their credit card accounts and to predict if a customer would cancel their accounts in the future.

In the past couple of years, the bank has experienced a record number of customers who have closed their credit card accounts and this has led to a decline in revenue. The bank's motive is to increase the number of credit lines to increase profitability.

The bank's goal is to improve its ability to detect customers who are at risk of having their accounts closed by which their financial losses would be minimized.

>Highlights and key findings:

From the bar plot, it is observed that customers with Part-time employment status have the maximum number of customers closing the credit card account. Uncertainty of job/income might be the reason that they can't afford a credit card and thus have closed their account. Full-time employees have the maximum count who has active credit card account, stable job and income might be the reason for active customer status.

Customers with divorced marital status are the least in the number who have active credit card accounts in comparison to single and married customers. It is also seen that customers with divorced marital status are the least in the number who have closed the credit card account. Customers who are single and married are found to be the highest who have active and closed account status. Hence, I found that there is no relation between marital status and customer status.

From the observation of the histogram, we find that active customers have a higher utilization percentage than those who have closed their accounts. This could mean that customers who don't pay using a credit card are willing to close their accounts.

People having education as associates and masters are the highest who have closed credit card accounts among all others. Similarly, associates and masters are also the highest among all who have the maximum active account status. Hence we can infer that education does not have a significant effect on customer status.

According to the Box plot, the median of active customers is slightly higher than the median of customers who have closed their accounts. Thus we can't conclude that customers with high salaries have active credit accounts because the difference is very small.


>Best classification model:

After, using different machine learning algorithms such as random forest, logistics regression, and KNN model, it is found that random forest gives the best performance in predicting this data. Important measures like ROC, F-measure, Recall and accuracy were used to conclude which is the best classification model. Random forest gives a high F-measure and accuracy of 0.95 and recall of 0.96 which is high than the rest of the models. Hence, Random Forest is a better classification model.

>Recommendations to the bank:

I would recommend the bank to provide more appealing offers like cashback on online shopping, dining, and grocery stores. By this, people might be tempted to use credit cards more. This would help the bank to retain the customers by not closing their accounts.

I would also recommend the bank to provide special schemes for customers with marital status as single and married. Because these are the customers who have the highest active and closed credit accounts. The bank should implement special policies to retain these customers so that it will enhance the bank's profit and customer base.
